---
title: "gNATSGO_burial_tool"
author: "Original scripts by Alex Wyckoff, compiled by Leah Clayton"
date: "2025-01-27"
output: html_document
---
Set working directories and output file names throughout before running
Ensure gNATSGO data is downloaded as "gNATSGO_Agg.csv"
Most relevant script appears to be (3), but these have not been tested

(1) Preliminary_Tool_Script_Updated.R
```{r}
### --- Burial Tool Script --- ###

# Load relevant packages
library(dplyr)
library(tidyr)

# Set working directory
setwd("/set_working_directory")

# Load aggregated gNATSGO table
gNATSGO_Agg <- read.csv("gNATSGO_Agg.csv")
head(gNATSGO_Agg)

# Calculate horizon thickness
gNATSGO_Agg <- gNATSGO_Agg %>%
  mutate(horizon_thickness_r = hzdepb_r - hzdept_r)

# Calculate total horizon thickness for each cokey
total_horizon_thickness <- gNATSGO_Agg %>%
  group_by(cokey) %>%
  summarise(total_horizon_thickness_r = sum(horizon_thickness_r))

# Calculate weighted averages for the relevant variables
weighted_averages <- gNATSGO_Agg %>%
  drop_na(claytotal_r, sandtotal_r, silttotal_r, awc_r) %>%
  group_by(cokey) %>%
  summarise(across(
    c(
      claytotal_r,
      sandtotal_r,
      silttotal_r,
      awc_r,
      om_r,
      ksat_r,
      aws025wta,
      aws050wta,
      aws0100wta,
      aws0150wta,
      wfifteenbar_r, 
      wthirdbar_r,
    ),
    ~ sum(. * horizon_thickness_r) / sum(horizon_thickness_r),
    .names = "weighted_{.col}"
  ))

# Combine the total horizon thickness and weighted averages
all_relevant_variables_weighted_means <-
  merge(total_horizon_thickness, weighted_averages, by = "cokey")%>%
  select(cokey, mukey, everything())

# Export for processing in ArcGIS
write.csv(all_relevant_variables_weighted_means, 
          file = 'output_location/all_relevant_variables_weighted_means.csv')
```

(2) gNATSGO Processing Script
```{r}
### --- Preliminary Burial Tool Script --- ###

# Load relevant packages
library(dplyr)
library(tidyr)

# Set working directory
setwd("/set_working_directory")

# Load aggregated gNATSGO table
gNATSGO_Agg <- read.csv("gNATSGO_Agg.csv")
head(gNATSGO_Agg)

# Calculate horizon thickness
gNATSGO_Agg <- gNATSGO_Agg %>%
  mutate(horizon_thickness_r = hzdepb_r - hzdept_r)

# Calculate total horizon thickness for each cokey
total_horizon_thickness <- gNATSGO_Agg %>%
  group_by(cokey) %>%
  summarise(total_horizon_thickness_r = sum(horizon_thickness_r))

# Calculate weighted averages for the relevant variables
weighted_averages <- gNATSGO_Agg %>%
  drop_na(claytotal_r, sandtotal_r, silttotal_r, awc_r) %>%
  group_by(cokey) %>%
  summarise(across(
    c(
      claytotal_r,
      sandtotal_r,
      silttotal_r,
      awc_r,
      om_r,
      ksat_r,
      aws025wta,
      aws050wta,
      aws0100wta,
      aws0150wta,
      wfifteenbar_r, 
      wthirdbar_r,
    ),
    ~ sum(. * horizon_thickness_r) / sum(horizon_thickness_r),
    .names = "weighted_{.col}"
  ))

# Combine the total horizon thickness and weighted averages
all_relevant_variables_weighted_means <-
  merge(total_horizon_thickness, weighted_averages, by = "cokey")%>%
  select(cokey, mukey, everything())

# Export for processing in ArcGIS
write.csv(all_relevant_variables_weighted_means, file = 'output_location/file_name.csv')
```

(3) gNATSGO_Processing_Script.R
```{r}
library(dplyr)
library(tidyr)

setwd('/set_working_directory')
gNATSGO_Agg <- read.csv("gNATSGO_Agg.csv")

# Calculate weighted averages by 'cokey'
weighted_avg_cokey <- gNATSGO_Agg %>%
  mutate(horizon_thickness = hzdepb_r - hzdept_r,
         total_thickness = sum(horizon_thickness)) %>%
  drop_na(awc_r) %>%
  group_by(cokey) %>%
  summarise(
    claytotal_r = weighted.mean(claytotal_r, w = horizon_thickness / total_thickness),
    sandtotal_r = weighted.mean(sandtotal_r, w = horizon_thickness / total_thickness),
    silttotal_r = weighted.mean(silttotal_r, w = horizon_thickness / total_thickness),
    om_r = weighted.mean(om_r, w = horizon_thickness / total_thickness),
    ksat_r = weighted.mean(ksat_r, w = horizon_thickness / total_thickness),
    wfifteenbar_r = weighted.mean(wfifteenbar_r, w = horizon_thickness / total_thickness),
    wthirdbar_r = weighted.mean(wthirdbar_r, w = horizon_thickness / total_thickness),
    awc_r = weighted.mean(awc_r, w = horizon_thickness / total_thickness)
  )

# Calculate weighted averages by 'mukey'
weighted_avg_mukey <- weighted_avg_cokey %>%
  inner_join(select(gNATSGO_Agg, mukey, cokey, comppct_r), by = "cokey") %>%
  drop_na(comppct_r) %>%
  group_by(mukey) %>%
  summarise(
    claytotal_r = weighted.mean(claytotal_r, w = comppct_r / 100),
    sandtotal_r = weighted.mean(sandtotal_r, w = comppct_r / 100),
    silttotal_r = weighted.mean(silttotal_r, w = comppct_r / 100),
    om_r = weighted.mean(om_r, w = comppct_r / 100),
    ksat_r = weighted.mean(ksat_r, w = comppct_r / 100),
    wfifteenbar_r = weighted.mean(wfifteenbar_r, w = comppct_r / 100),
    wthirdbar_r = weighted.mean(wthirdbar_r, w = comppct_r / 100),
    awc_r = weighted.mean(awc_r, w = comppct_r / 100)
  )

# Print rows to test dataset by hand
print_weighted_avg_mukey <- weighted_avg_mukey %>% filter(mukey == 507746)
print(print_weighted_avg_mukey)

# Print rows to test dataset by hand
print_gNATSGO_Agg <- gNATSGO_Agg %>% filter(mukey == 507710)
print(print_gNATSGO_Agg)

write.csv(weighted_avg_mukey, 
          file = '/output_location/Weighted_Avg_Mukey.csv')
```

(4) om_script.R -- should be included in (3)

om_r: A representative value of the amount by weight of decomposed plant and 
animal residue expressed as a weight percentage of the less than 2 mm soil material.
```{r}
library(dplyr)
library(tidyr)

setwd('/set_working_directory')
gNATSGO_Agg <- read.csv("gNATSGO_Agg.csv")

# Calculate weighted averages by 'cokey'
weighted_avg_cokey <- gNATSGO_Agg %>%
  mutate(horizon_thickness = hzdepb_r - hzdept_r,
         total_thickness = sum(horizon_thickness)) %>%
  drop_na(om_r) %>%
  group_by(cokey) %>%
  summarise(
    om_r = weighted.mean(om_r, w = horizon_thickness / total_thickness),
    
  )

# Calculate weighted averages by 'mukey'
weighted_avg_mukey <- weighted_avg_cokey %>%
  inner_join(select(gNATSGO_Agg, mukey, cokey, comppct_r), by = "cokey") %>%
  drop_na(comppct_r) %>%
  group_by(mukey) %>%
  summarise(
    om_r = weighted.mean(om_r, w = comppct_r / 100),
  )

write.csv(weighted_avg_mukey, 
          file = 'output_location/Weighted_Avg_Mukey_om_r.csv')
```

(5) kwfact_script.R

kwfact: An erodibility factor which quantifies the susceptibility of soil 
particles to detachment and movement by water. This factor is adjusted for the
effect of rock fragments.
```{r}
library(dplyr)
library(tidyr)

setwd('/set_working_directory')
gNATSGO_Agg <- read.csv("gNATSGO_Agg.csv")

# Calculate weighted averages by 'cokey'
weighted_avg_cokey <- gNATSGO_Agg %>%
  mutate(horizon_thickness = hzdepb_r - hzdept_r,
         total_thickness = sum(horizon_thickness)) %>%
  drop_na(kwfact) %>%
  group_by(cokey) %>%
  summarise(
    kwfact = weighted.mean(kwfact, w = horizon_thickness / total_thickness),
    
  )

# Calculate weighted averages by 'mukey'
weighted_avg_mukey <- weighted_avg_cokey %>%
  inner_join(select(gNATSGO_Agg, mukey, cokey, comppct_r), by = "cokey") %>%
  drop_na(comppct_r) %>%
  group_by(mukey) %>%
  summarise(
    kwfact = weighted.mean(kwfact, w = comppct_r / 100),
  )

write.csv(weighted_avg_mukey, 
          file = 'output_location/Weighted_Avg_Mukey_kwfact.csv')
```
